Contracts for Haskell

### 2011-07-05 contracts.tex

\begin{center}
\begin{array}{rclr}
  u,e &:=& x \mid f \mid \lambda x.~e \mid e~e \mid D(e,\dots,e) \mid \bad \mid n \mid e \mbox{ `satisfies` } c \mid \mbox{assert } (e,c) \mbox{ in } e  & (Expression)\\
  &\mid& \mbox{case } e \mbox{ of } [(p_i,e_i)] \mid \cf{e}&\\
  p &:=& \Delta,p \mid T,p \mid f \in c,p \mid \epsilon & (Program)\\
  \Delta &:=& d \mid opaque(d) & (Defintion~and~scope)\\
  d &:=& f~x_1 \dots x_n ::: c = e \mid f~x_1 \dots x_n ::: c = \mbox{ case } e \mbox{ of } [(pat_i,e_i)] & (Definition)\\
  T &:=& \mbox{data } x = DCons  & (Data~type~definition)\\ %TODO make type clearer
  DCons &:=& \epsilon \mid D ; DCons \mid D ::: c ; DCons & (List~of~data~constructors)\\
  pat &:=& D(x_1,\dots,x_n) & (Pattern)\\
\end{array}
\end{center}

@@

\begin{center}
\begin{array}{rclr}
  u,e &:=& x \mid f \mid e~e \mid D(e,\dots,e) \mid \bad \mid n \mid e \mbox{ `satisfies` } c \mid \cf{e} & (Expression)\\
  p &:=& \Delta,p \mid T,p \mid f \in c,p \mid \epsilon & (Program)\\
  \Delta &:=& d \mid opaque(d) & (Defintion~and~scope)\\
  d &:=& f~x_1 \dots x_n ::: c = e\mid f~x_1 \dots x_n ::: c = \mbox{ case } e \mbox{ of } [(pat_i,e_i)] & (Definition)\\
  T &:=& \mbox{data } x = DCons  & (Data~type~definition)\\ %TODO make type clearer
  DCons &:=& \epsilon \mid D ; DCons \mid D ::: c ; DCons & (List~of~data~constructors)\\
  pat &:=& D(x_1,\dots,x_n) & (Pattern)\\
\end{array}
\end{center}

@@

\subsection{FOL}
\begin{center}
\begin{array}{rclr}
  t &:=& x \mid \mbox{app}(t_1,t_2) \mid D(t,\dots,t) \mid f \mid n \mid \bad \mid \unr \mid \mbox{CF}(t) & (Term)\\
  \phi &:=& \forall x.\phi \mid \phi \to \phi \mid \lnot \phi \mid \phi \lor \phi \mid \phi \land \phi \mid true \mid t=t \mid \mbox{CF}(t)& (Formula)\\
\end{array}
\end{center}

@@

\begin{eqnarray}
\etrans{x} &=& (x,\emptyset)\\
\etrans{f} &=& (f,\emptyset)\\
\etrans{e_1~e_2} &=& (app(\etranst{e_1},\etranst{e_2}),\etransf{e_1} \cup \etransf{e_2})\\
\etrans{D(e_1,\dots,e_n)} &=& (D(\etranst{e_1},\dots,\etranst{e_n}),\cup_{1 \leq i \leq n} \etransf{e_i})\\
\etrans{\bad} &=& (\bad,\emptyset)\\
\etrans{n} &=& (n,\emptyset)\\
\etrans{e \mbox{ `satisfies` } c} &=& (\etranst{e},\etransf{e} \cup \forall x. \strans{x \in c} \leftrightarrow x~`satisfies`~c = True)\\
\etrans{\cf{e}} &=& \cf{e}\\
\end{eqnarray}

@@

\begin{eqnarray}
\dtrans{f~\overline{x} = e} &=& \forall x_1 \dots x_n.\etranst{f~x_1 \dots x_n} = \etranst{e} \cup \etransf{e}\\
\dtrans{f~\overline{x} = \mbox{case } e \mbox{ of } [D_i(\overline{z}) \mapsto e_i]} &=& \forall x_1 \dots x_n.(\bigwedge_i (\forall \overline{z}~\etranst{e} = \etranst{D_i(\overline{z})} \to \etranst{f~\overline{x}} = \etranst{e_i})\\
&& \land \etranst{e} = \bad \to \etranst{f~x_1 \dots x_n} = \bad)\\ 
&& \land ((\bigwedge_i e \neq D_i(sel_{1,D_i}(e),\dots,sel_{n_i,D_i}(e)) \land \etranst{e} \neq \bad)\\
&& ~~\to \etranst{f~x_1 \dots x_n} = \unr)\\
&& \cup \etransf{e}
\end{eqnarray}

@@

Equation (11) and (12) state that if we have a type mismatch than the code is $\unr$. It should never occur because we're assuming that the code has already been typechecked.

>>>>

\begin{center}
\begin{array}{rclr}
  u,e &:=& x \mid f \mid \lambda x.~e \mid e~e \mid D(e,\dots,e) \mid \bad \mid n \mid e \mbox{ `satisfies` } c \mid \mbox{assert } (e,c) \mbox{ in } e  & (Expression)\\
  &\mid& \mbox{case } e \mbox{ of } [(p_i,e_i)] \mid \cf{e} \mid f(e,\dots,e)&\\
  p &:=& \Delta,p \mid T,p \mid f \in c,p \mid \epsilon & (Program)\\
  \Delta &:=& d \mid opaque(d) & (Defintion~and~scope)\\
  d &:=& f~x_1 \dots x_n ::: c = e \mid f~x_1 \dots x_n ::: c = \mbox{ case } e \mbox{ of } [(pat_i,e_i)] & (Definition)\\
  T &:=& \mbox{data } x = DCons  & (Data~type~definition)\\ %TODO make type clearer
  DCons &:=& \epsilon \mid D ; DCons \mid D ::: c ; DCons & (List~of~data~constructors)\\
  pat &:=& D(x_1,\dots,x_n) & (Pattern)\\
\end{array}
\end{center}

@@

\begin{center}
\begin{array}{rclr}
  u,e &:=& x \mid f \mid e~e \mid D(e,\dots,e) \mid \bad \mid n \mid e \mbox{ `satisfies` } c \mid \cf{e} \mid f(e,\dots,e) & (Expression)\\
  p &:=& \Delta,p \mid T,p \mid f \in c,p \mid \epsilon & (Program)\\
  \Delta &:=& d \mid opaque(d) & (Defintion~and~scope)\\
  d &:=& f~x_1 \dots x_n ::: c = e\mid f~x_1 \dots x_n ::: c = \mbox{ case } e \mbox{ of } [(pat_i,e_i)] & (Definition)\\
  T &:=& \mbox{data } x = DCons  & (Data~type~definition)\\ %TODO make type clearer
  DCons &:=& \epsilon \mid D ; DCons \mid D ::: c ; DCons & (List~of~data~constructors)\\
  pat &:=& D(x_1,\dots,x_n) & (Pattern)\\
\end{array}
\end{center}

@@

\subsection{FOL}
\begin{center}
\begin{array}{rclr}
  t &:=& x \mid \mbox{app}(t_1,t_2) \mid D(t,\dots,t) \mid f(t,\dots,t) \mid f \mid n \mid \bad \mid \unr & (Term)\\
  \phi &:=& \forall x.\phi \mid \phi \to \phi \mid \lnot \phi \mid \phi \lor \phi \mid \phi \land \phi \mid true \mid t=t \mid \mbox{CF}(t)& (Formula)\\
\end{array}
\end{center}

@@

\begin{eqnarray}
\etrans{x} &=& (x,\emptyset)\\
\etrans{f} &=& (f,\emptyset)\\
\etrans{e_1~e_2} &=& (app(\etranst{e_1},\etranst{e_2}),\etransf{e_1} \cup \etransf{e_2})\\
\etrans{D(e_1,\dots,e_n)} &=& (D(\etranst{e_1},\dots,\etranst{e_n}),\cup_{1 \leq i \leq n} \etransf{e_i})\\
\etrans{\bad} &=& (\bad,\emptyset)\\
\etrans{n} &=& (n,\emptyset)\\
\etrans{e \mbox{ `satisfies` } c} &=& (\etranst{e},\etransf{e} \cup \forall x. \strans{x \in c} \leftrightarrow x~`satisfies`~c = True)\\
\etrans{f(e_1,\dots,e_n)} &=& f(\etranst{e_1},\dots,\etranst{e_n})\\
\etrans{\cf{e}} &=& (True,\cf{\eTranst{e}})\\
\end{eqnarray}

@@

\begin{eqnarray}
\dtrans{f~\overline{x} = e} &=& \forall x_1 \dots x_n.\etranst{f~x_1 \dots x_n} = \etranst{e} \cup \etransf{e}\\
&\cup& (\forall x_1 \dots x_n. (\bigwedge \cf{x_i} \rightarrow \cf{f(x_1,\dots,x_n)}) \leftrightarrow \cf{f\_ptr}\\
&\cup& (\forall x_1 \dots x_n. (f(x_1,\dots,x_n) = app(...app(f\_ptr,x_1)...,x_n)))\\
\dtrans{f~\overline{x} = \mbox{case } e \mbox{ of } [D_i(\overline{z}) \mapsto e_i]} &=& \forall x_1 \dots x_n.(\bigwedge_i (\forall \overline{z}~\etranst{e} = \etranst{D_i(\overline{z})} \to \etranst{f~\overline{x}} = \etranst{e_i})\\
&& \land \etranst{e} = \bad \to \etranst{f~x_1 \dots x_n} = \bad)\\ 
&& \land ((\bigwedge_i e \neq D_i(sel_{1,D_i}(e),\dots,sel_{n_i,D_i}(e)) \land \etranst{e} \neq \bad)\\
&& ~~\to \etranst{f~x_1 \dots x_n} = \unr)\\
&& \cup \etransf{e}\\
&\cup& (\forall x_1 \dots x_n. (\bigwedge \cf{x_i} \rightarrow \cf{f(x_1,\dots,x_n)}) \leftrightarrow \cf{f\_ptr}\\
&\cup& (\forall x_1 \dots x_n. (f(x_1,\dots,x_n) = app(...app(f\_ptr,x_1)...,x_n)))\\
\end{eqnarray}

@@

Equation (11) and (12) state that if we have a type mismatch than the code is $\unr$. It should never occur because we're assuming that the code has already been typechecked. %TODO eq number

Finally, we have to relate $f\_ptr$ to $f$ and define when we can say that $f$ is crash-free.
